Technology for application family creation based on domain analysis

### 2016-04-09 2016-domainAnalysis/domainAnalysis.tex
Remaining literature and some fixes

\author{
	\IEEEauthorblockN{Gudoshnikova Anna}
	\IEEEauthorblockA{
		Chair of informatics \\
		St.Petersburg State University (SPbSU) \\
		St.Petersburg, Russia \\
		Email: gudoshnikova.anna@gmail.com
	}
\and
	\IEEEauthorblockN{Yurii Litvinov}
	\IEEEauthorblockA{
    	Software Engineering chair \\
    	St.Petersburg State University (SPbSU) \\
    	St.Petersburg, Russia \\
    	Email: y.litvinov@spbu.ru 
	}
}

@@

\section*{Introduction}
The term ``reuse'' in software engineering closely associated with context. Reuse objects can be programs, parts of programs, specifications, requirements, architectures, test plans, etc. Reuse of one object leads to reuse of  another object. This means, there is a need to reuse something more than just code, i.e. there is a call for increasing the abstraction level. It is commonly supposed that reuse, as some kind of activity, can be divided into groups according to what should be reused: components, process for gaining the product, the technology or knowledge. At all accounts any reuse object cannot be discussed without environment, where the given object exists. Hence, the context problem is still remains.

@@

Despite different understanding the term ``domain analysis'' Arango [] showed that all formal domain analysis methods follow the general process for obtain the domain model. This process includes next stages: (1) domain characterization, (2) data collection, (3) data analysis, (4) classification and finally (5) evaluation of domain model. There are following domain analysis approaches: 1) DARE (Domain Analysis and Reuse Environment) []. The crucial idea of this method is creating domain book. This book include the universal architecture and library of reusable components. 2) DSSA (Domain-Specific Software Architectures) []. Given approach allows to create a domain glossary with the aid of use case analysis. 3) ODE (Ontology-based Domain Engineering) [] This approach connects the ontology idea with object-oriented approach. Ontology includes  terms and their connections, definitions, properties and constraints. Library of objects is built based on mapping ontology with object-oriented entities. 4) FODA (Feature-Oriented Domain Analysis)[]. This method has get popularity among scientists in the research area because of its simplicity for non-programmers. The main idea of this approach is creating feature model. This model describes functionality, which the future product should possess. Such model must note, what features are compulsory for implement in any instance of application in a given domain, what features must be implemented but there is some alternative between them, and present features, which may be implemented but not compulsory. This model can be easily built by expert in the domain.

Concerning product line creating with the aid of using domain model, Estublier[] presented approach, which is based on some aspects and requirements. These entities were proposed by authors. Such approach based on MDE methodology. Domain model is considered as  metamodel, which is described on MOF or UML. There is an interpreter, which translates each term in metamodel into Java-class, and concrete models --- into instances of these classes. Domain model is accompanied with feature model, which include some external behavior of the system. Authors use aspect-oriented techniques for feature implementing and following their  mapping with terms in domain model. Consequently, there is a close interaction between domain modeling and feature modeling. It seems that such approach is a bit complicated for non-programmers. In addition, there is no any industrial use of this method, but it is worth noting that authors describe appliance in this article [].

\begin{figure*}[ht]
	\centering
	\includegraphics[width=0.8\textwidth]{process.png}
	\caption{Relations between artefacts and roles in proposed approach to domain components reuse.}
	\label{image:process}
\end{figure*}

\section{Proposed approach}
\label{chapter:proposed}
In our approach we will use Feature-Oriented Domain Analysis (FODA) method to perform domain analysis and to create feature models. For this we will use visual editor that implements feature diagrams and is easy enough to use that domain experts can use it without problems. Then, when feature models are ready, each feature is implemented as reusable and configurable component on selected implementation language (C\#, C++, Java and so on) and feature library is formed as a collection of such components. This process requires qualified programmers and requires more effort than to simply create one application, but it allows to reuse features from feature library to create as many applications as needed. Also, this process is scalable, so we may add new features into feature library later, thus allowing to create more complex applications. At this stage of development domain experts shall work with programmers, and they shall use feature diagrams as an input for creation of feature library to simplify matching between features and components in feature library.

Next step is to create domain-specific language that allows to combine and configure features from feature library to implement applications in given domain. This is where our approach differs from common reuse strategies. Naive approach would be to generate an application directly from feature diagram, somehow marking features that shall be included into application, and it actually works fine when domain variability is low []. But more common is the situation when features themselves have properties that allow to configure them, those properties can have different types. Also, components may be related to each other in different ways, be used in configuration of one another, or some of their properties may be meaningless in absence or presence of other feature. Those rules may be implemented implicitly in application generator and require that programmers will always observe them, but we propose that these rules will be captured explicitly by dedicated domain-specific language. Such language may make models that do not observe those rules syntactically incorrect, and it will greatly reduce the possibility of human error and reduce knowledge required to efficiently use programming system.

>>>>

\author{
	\IEEEauthorblockN{Gudoshnikova Anna}
	\IEEEauthorblockA{
		Chair of informatics \\
		St.-Petersburg State University \\
		St.Petersburg, Russia \\
		Email: gudoshnikova.anna@gmail.com
	}
\and
	\IEEEauthorblockN{Yurii Litvinov}
	\IEEEauthorblockA{
    	Software Engineering chair \\
    	St.-Petersburg State University \\
    	St.Petersburg, Russia \\
    	Email: y.litvinov@spbu.ru 
	}
}

@@

\section{Introduction}
The term ``reuse'' in software engineering closely associated with context. Reuse objects can be programs, parts of programs, specifications, requirements, architectures, test plans, etc. Reuse of one object leads to reuse of  another object. This means, there is a need to reuse something more than just code, i.e. there is a call for increasing the abstraction level. It is commonly supposed that reuse, as some kind of activity, can be divided into groups according to what should be reused: components, process for gaining the product, the technology or knowledge. At all accounts any reuse object cannot be discussed without environment, where the given object exists. Hence, the context problem is still remains.

@@

Despite different understanding the term ``domain analysis'' Arango~\cite{arango1994domain} showed that all formal domain analysis methods follow the general process for obtain the domain model. This process includes next stages: (1) domain characterization, (2) data collection, (3) data analysis, (4) classification and finally (5) evaluation of domain model. There are following domain analysis approaches: 1) DARE (Domain Analysis and Reuse Environment)~\cite{frakes1998dare}. The crucial idea of this method is creating domain book. This book include the universal architecture and library of reusable components. 2) DSSA (Domain-Specific Software Architectures)~\cite{taylor1995software}. Given approach allows to create a domain glossary with the aid of use case analysis. 3) ODE (Ontology-based Domain Engineering)~\cite{falbo2002ontological}. This approach connects the ontology idea with object-oriented approach. Ontology includes  terms and their connections, definitions, properties and constraints. Library of objects is built based on mapping ontology with object-oriented entities. 4) FODA (Feature-Oriented Domain Analysis)~\cite{kang1990feature}. This method has get popularity among scientists in the research area because of its simplicity for non-programmers. The main idea of this approach is creating feature model. This model describes functionality, which the future product should possess. Such model must note, what features are compulsory for implement in any instance of application in a given domain, what features must be implemented but there is some alternative between them, and present features, which may be implemented but not compulsory. This model can be easily built by expert in the domain.

Concerning product line creating with the aid of using domain model, Estublier~\cite{estublier2005reuse} presented approach, which is based on some aspects and requirements. These entities were proposed by authors. Such approach based on MDE methodology. Domain model is considered as  metamodel, which is described on MOF or UML. There is an interpreter, which translates each term in metamodel into Java-class, and concrete models --- into instances of these classes. Domain model is accompanied with feature model, which include some external behavior of the system. Authors use aspect-oriented techniques for feature implementing and following their  mapping with terms in domain model. Consequently, there is a close interaction between domain modeling and feature modeling. It seems that such approach is a bit complicated for non-programmers. In addition, there is no any industrial use of this method, but it is worth noting that authors describe appliance in this article~\cite{estublier2003approach}.

\begin{figure*}[ht]
	\centering
	\includegraphics[width=0.76\textwidth]{process.png}
	\caption{Relations between artefacts and roles in proposed approach to domain components reuse.}
	\label{image:process}
\end{figure*}

\section{Proposed approach}
\label{chapter:proposed}
In our approach we will use some ideas of Feature-Oriented Domain Analysis (FODA) method to perform domain analysis and to create feature models. For this we will use visual editor that implements feature diagrams and is easy enough for domain experts. Then, when feature models are ready, each feature is implemented as reusable and configurable component on selected implementation language (C\#, C++, Java and so on) and feature library is formed as a collection of such components. This process requires qualified programmers and requires more effort than to simply create one application, but it allows to reuse features from feature library to create as many applications as needed. Also, this process is scalable, so we may add new features into feature library later, thus allowing to create more complex applications. At this stage of development domain experts shall work with programmers, and they shall use feature diagrams as an input for creation of feature library to simplify matching between features and components in feature library.

Next step is to create domain-specific language that allows to combine and configure features from feature library to implement applications in given domain. This is where our approach differs from common reuse strategies. Naive approach would be to generate an application directly from feature diagram, somehow marking features that shall be included into application, and it actually works fine when domain variability is low. But more common is the situation when features themselves have properties that allow to configure them, those properties can have different types. Also, components may be related to each other in different ways, be used in configuration of one another, or some of their properties may be meaningless in absence or presence of other feature. Those rules may be implemented implicitly in application generator and require that programmers will always observe them, but we propose that these rules will be captured explicitly by dedicated domain-specific language. Such language may make models that do not observe those rules syntactically incorrect, and it will greatly reduce the possibility of human error and reduce knowledge required to efficiently use programming system.
